### 2025-09-23 paper/sec/authoring-workflow.tex
wow this is hard.

\item \textbf{Initial configuration.} Author imports text and accompanying data into the system, creating a
programmatic representation of the target document. Initially this representation is isomorphic to the target
text, namely a string literal of the form \kw{"""..."""}, where the triple quotes are Fluid syntax for an
\emph{interpolated string}, i.e.~a string literal where expressions of the form \kw{\{$e$\}} are permitted
within the string. The \SuggestionAgent analyses these inputs and identifies any fragments of natural language
which are candidates for being computed instead of remaining as literal substrings.

\item \textbf{High-level Authoring workflow.} The system then enters the human-in-the loop authoring workflow
shown in Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The
system initially waits for the author to select a fragment of text $s$ to interpret, a fragment which may or
may not have been highlighted previously by the \SuggestionAgent. The system then generates a candidate Fluid
expression $e$ using a closed-loop synthesis step ((3) below). If this process succeeds, the system replaces
the selected substring $s$ with the interpolation expression \kw{\{$e$\}}, creating a new document
configuration. This is the primary successful path through the workflow and returns the system to the entry
state where it is waiting for another top-level interaction from the author.

\item \textbf{Code synthesis step.} The expression synthesis step is itself an internal loop which starts with
an initial prompt sent to the LLM (\secref{prompt-design} below) requesting the generation of an expression
$e$. Using the Fluid command-line interface, the expression is validated by checking that it evaluates without
error, produces a value coercible to a string $s'$, and finally that $s'$ is equal to the target fragment $s$.
Any failure triggers prompt augmentation with the appropriate error message and the system retries generation.
If code synthesis loop is able to yield an expression which computes $s$ within a maximum
number of retries, the system proceeds to the manual validation step (4). Otherwise, the code synthesis step has failed and there are two possibilities:
\begin{enumerate}
\item If the last generated $e$ was invalid (resulting in an error), no remedial action is open to the author. This is considered an unsuccessful path through the workflow and returns the system to the start state.
\item In the $s' \neq s$ case, the user can choose to manually abort and return to the entry state, or
optionally to \emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the
candidate expression. This is intended to cover the situation where the author has made a claim which is
\emph{incorrect}, but the data set and surrounding natural language has caused the LLM to synthesise an
expression which generates a different value from the one specified by the user.
\end{enumerate}

\item \textbf{Manual validation step.} Once a candidate expression has been generated, the author is able to
republish the web page containing the Fluid article and interact with the fragment of text. As shown in
\secref{}, this is an important validation step that can reveal errors in the generated expression.

>>>>

\item \textbf{Initial configuration.} The author imports the target text and accompanying data into the
system, creating a programmatic representation of the target document. Initially this is simply isomorphic to
the target text, taking the form of a string literal \kw{"""..."""}, where the triple quotes are Fluid syntax
for a Python or JavaScript-style \emph{interpolated string}, i.e.~a literal where expressions of the form
\kw{\{$e$\}} are permitted within the string. The \SuggestionAgent analyses these inputs and identifies any
fragments of natural language which are candidates for being computed instead of remaining as literal
substrings.

\item \textbf{High-level Authoring workflow.} The system then enters the human-in-the loop authoring workflow
shown in Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The
system initially waits for the author to select a fragment of text $s$ to interpret (perhaps but not
necessarily highlighted by the \SuggestionAgent). The system then attempts to generates a candidate Fluid
expression $e$ using the closed-loop synthesis step (3) below. If this process succeeds with an expression
$e$, the system proceeds to the manual validation step (4) below. If the code generation step fails with no
expression, no remedial action is open to the author; this is considered an unsuccessful path through the
workflow and returns the system to the start state. Otherwise the code generation step fails with a mismatched
string $s' \neq s$ outcome, and the user can choose to manually abort and return to the entry state, or
optionally to \emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the
candidate expression. This is intended to cover the situation where the author has made a claim which is
\emph{incorrect}, but the data set and surrounding natural language has caused the LLM to synthesise an
expression which generates a different value from the one specified by the user.

\item \textbf{Code synthesis step.} The expression synthesis step is itself an internal loop which starts with
an initial prompt sent to the LLM (\secref{prompt-design} below) requesting the generation of an expression
$e$. Using the Fluid command-line interface, the expression is validated by checking that it evaluates without
error, produces a value coercible to a string $s'$, and finally that $s'$ is equal to the target fragment $s$.
Any failure triggers prompt augmentation with the appropriate error message and the system retries generation.
If code synthesis loop is able to yield an expression which computes $s$ within a maximum number of retries,
the synthesis step succeeds with $e$. If the last generated $e$ was invalid (resulting in an error), the code
synthesis step fails with no expression. Otherwise, the code synthesis fails with a mismatched string outcome.

\item \textbf{Manual validation step.} Once a candidate expression has been generated, the author is able to
republish the web page containing the Fluid article and interact with the fragment of text.  As shown in
\secref{}, this is an important validation step that can reveal errors in the generated expression. If the
interactions look reasonable, the author can approve the replacement of the selected substring $s$ with the
interpolation expression \kw{\{$e$\}}, creating a new document configuration. This is the primary successful
path through the workflow and returns the system to the entry state where it is waiting for another top-level
interaction from the author. Otherwise, the author can reject the proposed change and return to the start
state.
